\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2019-07-08 slides/Jonathan-Proof-Reconstruction-AIM.tex
[ slides ] fixed compilation

\newcommand{\gobackreconstruction}{
  \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.8cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = reconstruction]
      (go) at (0,0) {\tiny Go Back};
  \end{tikzpicture}}

@@

% --------------------------------------------------------------------------
% Beamer configuration.
% --------------------------------------------------------------------------
\usetheme{default}
\usecolortheme{default}
\usefonttheme{serif}

@@

% footer.
\setbeamercolor{headFoot}{fg=white, bg=plum!80!black}
\setbeamertemplate{footline}{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}
    [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
    \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|
    \hspace*{2mm}\textbf\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}
    [wd=.2\paperwidth,ht=2.3ex,dp=1ex,right]{headFoot}%
    \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
  \end{beamercolorbox}}%
  \vskip 0pt%
}

@@

% --------------------------------------------------------------------------
% Fonts
% --------------------------------------------------------------------------
\usefonttheme{professionalfonts}
\usefonttheme{serif}
\usepackage{fontspec}
\usepackage{mathtools}
\usepackage{unicode-math}

\newfontfamily\djvu[ExternalLocation=fonts/
  , BoldFont=DejaVuSansMono-Bold.ttf
  , BoldItalicFont=DejaVuSansMono-BoldOblique.ttf
  , ItalicFont=DejaVuSansMono-Oblique.ttf
  ]{DejaVuSansMono.ttf}

\setmonofont[ExternalLocation=fonts/
, BoldFont=DejaVuSansMono-Bold.ttf
, BoldItalicFont=DejaVuSansMono-BoldOblique.ttf
, ItalicFont=DejaVuSansMono-Oblique.ttf
]{DejaVuSansMono.ttf}

\setmathfont[ExternalLocation=fonts/
  ]{DejaVuMathTeXGyre.ttf}
\newfontfamily\mathfont{fonts/DejaVuMathTeXGyre.ttf}

\setmainfont[ExternalLocation=fonts/
  , BoldFont=SourceSansPro-Semibold.otf
  , BoldItalicFont=SourceSansPro-SemiboldIt.otf
  , ItalicFont=SourceSansPro-It.otf
  ]{SourceSansPro-Regular.otf}

\newfontfamily\sourcecode[ExternalLocation=fonts/
  , BoldFont=SourceCodePro-Semibold.ttf
  , BoldItalicFont=SourceCodePro-SemiboldIt.ttf
  , ItalicFont=SourceCodePro-It.ttf
  ]{SourceCodePro-Regular.ttf}

\usepackage{minted}
\setminted[cagda]{
  bgcolor   = aliceblue
, fontsize  = \footnotesize
, frame     = none
% , framerule = 0.4pt
% , framesep  = 0pt
, style     = cagda
}

@@

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}.}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] formula in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{itemize}
  \item $p ⊢ p$
\begin{tptp}
fof(myaxiom, axiom, p).
fof(goal, conjecture, p).
\end{tptp}

  \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
  \end{itemize}
\gobackreconstruction
\end{frame}

@@

\begin{center}
  {\footnotesize
\begin{tptp}
language(name, role, formula, source [,useful info]).
\end{tptp}}
\end{center}

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents).
\end{tptp}
}
\end{center}
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\end{frame}

@@

Is a \texttt{Haskell} program
that translates proofs given by \texttt{Metis}
in \texttt{TSTP} format to \texttt{Agda} code\\

@@

\begin{minted}{cagda}
data Prop : Set where
  Var : Fin n → Prop
  ⊤   : Prop
  ⊥   : Prop
  _∧_ : (φ ψ : Prop) → Prop
  _∨_ : (φ ψ : Prop) → Prop
  _⇒_ : (φ ψ : Prop) → Prop
  _⇔_ : (φ ψ : Prop) → Prop
  ¬_  : (φ : Prop)   → Prop
\end{minted}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{\texttt{Agda-Prop} Library~\citep{AgdaProp}}
\begin{itemize}
  \item A data type for theorems
  \vspace*{1mm}
\begin{minted}{cagda}
data _⊢_ : (Γ : Ctxt)(φ : Prop) → Set
\end{minted}

@@

\begin{minted}{cagda}
assume, axiom, weaken, ⊤-intro, ⊥-elim, ¬-intro,
¬-elim, ∧-intro, ∧-proj₁, ∧-proj₂, ∨-intro₁,
∨-intro₂, ∨-elim, ⇒-intro, ⇒-elim, ⇔-intro,
⇔-elim₁, ⇔-elim₂.
\end{minted}
  \item Natural deduction proofs for more than 71 theorems
  \vspace*{1mm}
\begin{minted}{cagda}
⇔-equiv, ⇔-assoc, ⇔-comm, ⇒-⇔-¬∨, ⇔-¬-to-¬,
¬⇔-to-¬, ¬¬-equiv, ⇒⇒-⇔-∧⇒, ⇔-trans, ∧-assoc,
∧-comm, ∧-dist, ¬∧-to-¬∨¬, ¬∨¬-to-¬∧, ¬∨¬-⇔-¬∧,
subst⊢∧₁, subst⊢∧₂, ∨-assoc, ∨-comm, ∨-dist,
∨-equiv, ¬∨-to-¬∧¬, ¬∧¬-to-¬∨, ∨-dmorgan,
¬¬∨¬¬-to-∨, cnf, nnf, dnf, RAA, ...
\end{minted}
\end{itemize}
\end{frame}

@@

\begin{minted}{cagda}
data ConjView : Prop → Set where
  conj  : (φ₁ φ₂ : Prop) → ConjView (φ₁ ∧ φ₂)
  other : (φ : Prop)     → ConjView φ

@@

conjunct-path : (φ ψ : Prop) → Path → Path
conjunct-path φ ψ path with ⌊ eq φ ψ ⌋
... | true  = path ∷ʳ pick
... | false with conj-view φ
...   | other _ = []
...   | conj φ₁ φ₂ with conjunct-path φ₁ ψ []
...     | subpath@(_ ∷ _) = (path ∷ʳ proj₁) ++ subpath
...     | [] with conjunct-path φ₂ ψ []
...          | subpath@(_ ∷ _) = (path ∷ʳ proj₂) ++ subpath
...          | []              = []
\end{minted}

\begin{frame}[fragile, label=atp-conjunct-proof]{The \texttt{conjunct} function and its theorem, \texttt{atp-conjunct}}
\vfill
\begin{minted}{cagda}
conjunct : Prop → Prop → Prop
conjunct φ ψ with conj-view φ | conjunct-path φ ψ []
... | _          | []        = φ
... | conj _ _   | pick  ∷ _ = φ
... | conj φ₁ _  | proj₁ ∷ _ = conjunct φ₁ ψ
... | conj _ φ₂  | proj₂ ∷ _ = conjunct φ₂ ψ
... | other .φ   | _         = φ

@@

atp-conjunct {Γ} {φ} ψ Γ⊢φ
  with conj-view φ | conjunct-path φ ψ []
...  | _           | []        = Γ⊢φ
...  | conj _ _    | pick  ∷ _ = Γ⊢φ
...  | conj _ _    | proj₁ ∷ _ = atp-conjunct ψ (∧-proj₁ Γ⊢φ)
...  | conj _ _    | proj₂ ∷ _ = atp-conjunct ψ (∧-proj₂ Γ⊢φ)
...  | other _     | (_ ∷ _)   = Γ⊢φ
\end{minted}
\vfill
\gobackreconstruction
\end{frame}

@@

\begin{itemize}
  \item The problem is $p ∧ q ⊢ q ∧ p$
  \item In \texttt{TPTP} format
\begin{tptp}
$ cat problem.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
  \item How to use Athena with your problem

\begin{tptp}
$ metis --show proof problem.tptp > problem.tstp
$ athena problem.tstp
$ agda problem.tstp
\end{tptp}
\end{itemize}
\end{frame}

\begin{frame}[fragile, allowframebreaks]{\texttt{TSTP} Derivation given by \texttt{Metis}}
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
fof(subgoal_0, plain, q,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, q => p,
    inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ q,
inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, (~ q),
    inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p & q,
    inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, q,
    inference(conjunct, [], [normalize_0_1])).
fof(normalize_0_3, plain, $false,
    inference(simplify, [],
        [normalize_0_0, normalize_0_2])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_3])).
fof(negate_1_0, plain, ~ (q => p),
    inference(negate, [], [subgoal_1])).
fof(normalize_1_0, plain, ~ p & q,
    inference(canonicalize, [], [negate_1_0])).
fof(normalize_1_1, plain, p & q,
     inference(canonicalize, [], [a])).
fof(normalize_1_2, plain, p,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_3, plain, q,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_4, plain, $false,
    inference(simplify, [],
      [normalize_1_0, normalize_1_2, normalize_1_3])).
cnf(refute_1_0, plain, ($false),
    inference(canonicalize, [], [normalize_1_4])).
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=verified-example]{Problem $p ∧ q ⊢ q ∧ p$}{Proof generated with \texttt{Athena}}
\vfill
\begin{minted}{cagda}
p, q, a, goal, subgoal₀, subgoal₁ : Prop

@@

\begin{frame}[fragile, label=verified-example-2]{Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
\begin{minted}{cagda}
a : Prop
a = p ∧ q

@@

proof₀ : Γ ⊢ subgoal₀
proof₀ =
  (RAA
    (atp-simplify ⊥
      (assume {Γ = Γ}
        (¬ subgoal₀))
      (atp-conjunct q
        (atp-canonicalize (p ∧ q)
          (weaken (¬ subgoal₀)
            (assume {Γ = ∅} a))))))
\end{minted}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-3]
  {Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
\begin{minted}{cagda}
subgoal₁ : Prop
subgoal₁ = q ⇒ p

proof₁ : Γ ⊢ subgoal₁
proof₁ =
  (RAA
    (atp-simplify ⊥
      (atp-conjunct q
        (atp-canonicalize (p ∧ q)
          (weaken (¬ subgoal₁)
            (assume {Γ = ∅} a))))
      (atp-simplify ⊥
        (atp-canonicalize ((¬ p) ∧ q)
          (assume {Γ = Γ}
            (¬ subgoal₁)))
        (atp-conjunct p
          (atp-canonicalize (p ∧ q)
            (weaken (¬ subgoal₁)
              (assume {Γ = ∅} a)))))))
\end{minted}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-4]{Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
\begin{minted}{cagda}
-- Premise.
Γ = [ a ]

@@

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal            -- q ∧ (q ⇒ p) ⇒ p
    (∧-intro proof₀ proof₁)
\end{minted}
\vfill
\gobackreconstruction
\end{frame}

>>>>

\newcommand{\gobackreconstruction}{
  \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.8cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = reconstruction]
      (go) at (0,0) {\tiny back};
  \end{tikzpicture}}

@@

% --------------------------------------------------------------------------
% Beamer configuration.
% --------------------------------------------------------------------------
\usetheme{default}
\usecolortheme{default}
% \usefonttheme{serif}

@@

% footer.
\setbeamercolor{headFoot}{fg=white, bg=plum!80!black}
\setbeamertemplate{footline}{
  \leavevmode%
  \hbox{%
  % \begin{beamercolorbox}
  %   [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
  %   \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|
  %   \hspace*{2mm}\textbf\insertshortauthor
  % \end{beamercolorbox}%
  \begin{beamercolorbox}
    [wd=\paperwidth,ht=2.4ex,dp=1ex,right]{headFoot}%
    \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
  \end{beamercolorbox}}%
  \vskip 0pt%
}

% \setbeamertemplate{footline}{
%   \leavevmode%
%   \hbox{%
%   \begin{beamercolorbox}
%     [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
%     \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|
%     \hspace*{2mm}\textbf\insertshortauthor
%   \end{beamercolorbox}%
%   \begin{beamercolorbox}
%     [wd=.2\paperwidth,ht=2.3ex,dp=1ex,right]{headFoot}%
%     \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
%   \end{beamercolorbox}}%
%   \vskip 0pt%
% }

@@

% --------------------------------------------------------------------------
% Fonts
% --------------------------------------------------------------------------
\usefonttheme{professionalfonts}
% \usefonttheme{serif}

\usepackage{fontspec}
\usepackage{mathtools}
\usepackage{unicode-math}

\usepackage{minted}

@@

\renewcommand*\rmdefault{cmr}

% -------------------------------------------------------------------
% New commands
% -------------------------------------------------------------------

\usepackage{textcomp}
\usepackage[verbose]{newunicodechar}

\newunicodechar{¬}{\ensuremath{\neg}\;}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{γ}{\ensuremath{\gamma}}
\newunicodechar{φ}{\ensuremath{\varphi}}
\newunicodechar{ψ}{\ensuremath{\psi}}
\newunicodechar{ϕ}{\ensuremath{\varphi}}
\newunicodechar{ᵢ}{\ensuremath{{}_{i}}}
\newunicodechar{₀}{\ensuremath{{}_{0}}}
\newunicodechar{₁}{\ensuremath{{}_{1}}}
\newunicodechar{₂}{\ensuremath{{}_{2}}}
\newunicodechar{₃}{\ensuremath{{}_{3}}}
\newunicodechar{₄}{\ensuremath{{}_{4}}}
\newunicodechar{₅}{\ensuremath{{}_{5}}}
\newunicodechar{₆}{\ensuremath{{}_{6}}}
\newunicodechar{₇}{\ensuremath{{}_{7}}}
\newunicodechar{₈}{\ensuremath{{}_{8}}}
\newunicodechar{₉}{\ensuremath{{}_{9}}}
\newunicodechar{ₙ}{\ensuremath{{}_{n}}}
\newunicodechar{ₘ}{\ensuremath{{}_{m}}}
\newunicodechar{ℓ}{\ensuremath{\ell}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{⇒}{\ensuremath{\supset}}
\newunicodechar{⊃}{\ensuremath{\supset}}
\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{∘}{\ensuremath{\circ}}
\newunicodechar{∙}{\ensuremath{\bullet}}
\newunicodechar{∧}{\ensuremath{\wedge}}
\newunicodechar{∨}{\ensuremath{\vee}}
\newunicodechar{∼}{\ensuremath{\sim}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{⊕}{\ensuremath{\oplus}}
\newunicodechar{⊖}{\ensuremath{\ominus}}
\newunicodechar{⊢}{\ensuremath{⊢}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{⊻}{\ensuremath{\veebar}}
\newunicodechar{⟝}{\ensuremath{⊢}}
\newunicodechar{⬓}{\ensuremath{\square}}
\newunicodechar{ⱼ}{\ensuremath{{}_{j}}}
\newunicodechar{λ}{\ensuremath{\lambda}}

\usepackage{xspace}

\newcommand{\abbre}[1]{\textsc{#1}\xspace}
\newcommand{\ATPs}{\abbre{ATPs}}
\newcommand{\ATP}{\abbre{ATP}}
\newcommand{\Bool}{\abbre{Bool}}
\newcommand{\CNF}{\abbre{CNF}}
\newcommand{\CPL}{\abbre{CPL}}
\newcommand{\ITP}{\abbre{ITP}}
\newcommand{\ITPs}{\abbre{ITPs}}
\newcommand{\List}{\abbre{List}}
\newcommand{\SAT}{\abbre{SAT}}
\newcommand{\SMT}{\abbre{SMT}}

\newcommand{\name}[1]{\texttt{#1}\xspace}
\newcommand{\canonicalize}{\name{canonicalize}}
\newcommand{\clausify}{\name{clausify}}
\newcommand{\conjunct}{\name{conjunct}}
\newcommand{\negate}{\name{negate}}
\newcommand{\resolve}{\name{resolve}}
\newcommand{\simplify}{\name{simplify}}
\newcommand{\strip}{\name{strip}}
\newcommand{\tptpaxiom}{\name{axiom}}
\newcommand{\tptpconjecture}{\name{conjecture}}

\newcommand{\nnf}{\name{nnf}}

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Athena}{\prg{Athena}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}
\newcommand{\Metis}{\prg{Metis}}

\newcommand{\len}[1]{\texttt{#1}\xspace}
\newcommand{\Haskell}{\len{Haskell}}
\newcommand{\TPTP}{\len{TPTP}}
\newcommand{\TSTP}{\len{TSTP}}

% -- Type notation.
\newcommand{\type}[1]{ {\fontseries{sb}\selectfont \textsf{#1}} \xspace}
\newcommand{\NProp}{\type{Prop}}
\newcommand{\Nat}{\type{Nat}}
\newcommand{\Prop}{\type{Prop}}
\newcommand{\Source}{\type{Premise}}
\newcommand{\Target}{\type{Conclusion}}
\newcommand{\Lit}{\type{Lit}}
% --

\newcommand{\NN}{\ensuremath{\mathbb{N}}}
\newcommand{\fun}[1]{ {\fontseries{sb}\selectfont \textsf{#1}} \xspace}

\newcommand{\zero}{\fun{zero}}
\newcommand{\suc}{\fun{succ}}
\newcommand{\fassoc}{\fun{assoc}}
\newcommand{\fbuild}{\fun{build}}
\newcommand{\fcanonicalize}{\fun{canonicalize}}
\newcommand{\fcanon}{\fun{simplify}}
\newcommand{\fclausify}{\fun{clausify}}
\newcommand{\fcnf}{\fun{cnf}}
\newcommand{\fconjunctor}{\fun{disj}}
\newcommand{\fconjunct}{\fun{conjunct}}
\newcommand{\fdist}{\fun{dist}}
\newcommand{\ffactor}{\fun{factor}}
\newcommand{\fform}{\fun{form}}
\newcommand{\fmax}{\fun{max}}
\newcommand{\fndisj}{\fun{ndisj}}
\newcommand{\fnegate}{\fun{negate}}
\newcommand{\fnnfp}{\fun{nnf}₀}
\newcommand{\fnnf}{\fun{nnf}}
\newcommand{\fpolarity}{\fun{polarity}}
\newcommand{\frank}{\fun{rank}}
\newcommand{\frconj}{\fun{rconj}}
\newcommand{\frdisj}{\fun{rdisj}}
\newcommand{\freorder}{\fun{reorder}}
\newcommand{\fresolve}{\fun{resolve}}
\newcommand{\frm}{\fun{rm}}
\newcommand{\frsol}{\fun{rsol}}
\newcommand{\fsbuild}{\fun{sbuild}}
\newcommand{\fsimplify}{\fun{simplify}}
\newcommand{\fstripp}{\fun{strip}\prime}
\newcommand{\fstrip}{\fun{strip}}
\newcommand{\fsubst}{\fun{subst}}
\newcommand{\fuh}{\fun{uh}}

% latin etc. abbrev
\newcommand{\abbrev}[1]{#1} % alternative: \emph{#1}
\newcommand{\cf}{\abbrev{cf.}\ }
\newcommand{\eg}{\abbrev{e.\,g.}}
\newcommand{\Eg}{\abbrev{E.\,g.}}
\newcommand{\ie}{\abbrev{i.\,e.}}
\newcommand{\Ie}{\abbrev{I.\,e.}}
\newcommand{\etal}{\abbrev{et.\,al.}}
\newcommand{\wwlog}{w.\,l.\,o.\,g.} % \wlog is ``write into log file''
\newcommand{\Wlog}{W.\,l.\,o.\,g.}
\newcommand{\wrt}{w.\,r.\,t.}
\newcommand{\TheEnd}{\\\hspace*{0.975\textwidth}$\blacksquare$}

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}.}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}

\begin{verbatim}
language(name, role, formula).
\end{verbatim}

    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] formula in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{itemize}
  \item $p ⊢ p$
\begin{verbatim}
fof(myaxiom, axiom, p).
fof(goal, conjecture, p).
\end{verbatim}

  \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{verbatim}
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{verbatim}
  \end{itemize}
\gobackreconstruction
\end{frame}

@@

\begin{center}
  {\footnotesize
\begin{verbatim}
language(name, role, formula, source [,useful info]).
\end{verbatim}}
\end{center}

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{verbatim}
inference(rule, useful info, parents).
\end{verbatim}
}
\end{center}
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} for the problem $p ⊢ p$

{\footnotesize
\begin{verbatim}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{verbatim}
}
\end{itemize}
\end{frame}

@@

Is a \texttt{agda} program
that translates proofs given by \texttt{Metis}
in \texttt{TSTP} format to \texttt{Agda} code\\

@@

\begin{minted}{agda}
data Prop : Set where
  Var : Fin n → Prop
  ⊤   : Prop
  ⊥   : Prop
  _∧_ : (φ ψ : Prop) → Prop
  _∨_ : (φ ψ : Prop) → Prop
  _⇒_ : (φ ψ : Prop) → Prop
  _⇔_ : (φ ψ : Prop) → Prop
  ¬_  : (φ : Prop)   → Prop
\end{minted}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{\texttt{Agda-Prop} Library~\citep{AgdaProp}}
\begin{itemize}
  \item A data type for theorems
  \vspace*{1mm}
\begin{minted}{agda}
data _⊢_ : (Γ : Ctxt)(φ : Prop) → Set
\end{minted}

@@

\begin{minted}{agda}
assume, axiom, weaken, ⊤-intro, ⊥-elim, ¬-intro,
¬-elim, ∧-intro, ∧-proj₁, ∧-proj₂, ∨-intro₁,
∨-intro₂, ∨-elim, ⇒-intro, ⇒-elim, ⇔-intro,
⇔-elim₁, ⇔-elim₂.
\end{minted}
  \item Natural deduction proofs for more than 71 theorems
  \vspace*{1mm}
\begin{minted}{agda}
⇔-equiv, ⇔-assoc, ⇔-comm, ⇒-⇔-¬∨, ⇔-¬-to-¬,
¬⇔-to-¬, ¬¬-equiv, ⇒⇒-⇔-∧⇒, ⇔-trans, ∧-assoc,
∧-comm, ∧-dist, ¬∧-to-¬∨¬, ¬∨¬-to-¬∧, ¬∨¬-⇔-¬∧,
subst⊢∧₁, subst⊢∧₂, ∨-assoc, ∨-comm, ∨-dist,
∨-equiv, ¬∨-to-¬∧¬, ¬∧¬-to-¬∨, ∨-dmorgan,
¬¬∨¬¬-to-∨, cnf, nnf, dnf, RAA, ...
\end{minted}
\end{itemize}
\end{frame}

@@

{\footnotesize
\begin{minted}{agda}
data ConjView : Prop → Set where
  conj  : (φ₁ φ₂ : Prop) → ConjView (φ₁ ∧ φ₂)
  other : (φ : Prop)     → ConjView φ

@@

conjunct-path : (φ ψ : Prop) → Path → Path
conjunct-path φ ψ path with ⌊ eq φ ψ ⌋
... | true  = path ∷ʳ pick
... | false 
    with conj-view φ
...    | other _ = []
...    | conj φ₁ φ₂ with conjunct-path φ₁ ψ []
...       | subpath@(_ ∷ _) = (path ∷ʳ proj₁) ++ subpath
...       | [] with conjunct-path φ₂ ψ []
...          | subpath@(_ ∷ _) = (path ∷ʳ proj₂) ++ subpath
...          | []              = []
\end{minted}
}
\end{frame}

\begin{frame}[fragile, label=atp-conjunct-proof]{The \texttt{conjunct} function and its theorem, \texttt{atp-conjunct}}
\vfill
{\footnotesize
\begin{minted}{agda}
conjunct : Prop → Prop → Prop
conjunct φ ψ with conj-view φ | conjunct-path φ ψ []
... | _          | []        = φ
... | conj _ _   | pick  ∷ _ = φ
... | conj φ₁ _  | proj₁ ∷ _ = conjunct φ₁ ψ
... | conj _ φ₂  | proj₂ ∷ _ = conjunct φ₂ ψ
... | other .φ   | _         = φ

@@

atp-conjunct {Γ} {φ} ψ Γ⊢φ
  with conj-view φ | conjunct-path φ ψ []
...  | _           | []        = Γ⊢φ
...  | conj _ _    | pick  ∷ _ = Γ⊢φ
...  | conj _ _    | proj₁ ∷ _ = atp-conjunct ψ (∧-proj₁ Γ⊢φ)
...  | conj _ _    | proj₂ ∷ _ = atp-conjunct ψ (∧-proj₂ Γ⊢φ)
...  | other _     | (_ ∷ _)   = Γ⊢φ
\end{minted}
}
\vfill
\gobackreconstruction
\end{frame}

@@

\begin{itemize}
  \item The problem is $p ∧ q ⊢ q ∧ p$
  \item In \texttt{TPTP} format
\begin{verbatim}
$ cat problem.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{verbatim}
  \item How to use Athena with your problem

\begin{verbatim}
$ metis --show proof problem.tptp > problem.tstp
$ athena problem.tstp
$ agda problem.tstp
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile, allowframebreaks]{\texttt{TSTP} Derivation given by \texttt{Metis}}

\begin{verbatim}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
fof(subgoal_0, plain, q,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, q => p,
    inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ q,
inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, (~ q),
    inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p & q,
    inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, q,
    inference(conjunct, [], [normalize_0_1])).
fof(normalize_0_3, plain, $false,
    inference(simplify, [],
        [normalize_0_0, normalize_0_2])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_3])).
fof(negate_1_0, plain, ~ (q => p),
    inference(negate, [], [subgoal_1])).
fof(normalize_1_0, plain, ~ p & q,
    inference(canonicalize, [], [negate_1_0])).
fof(normalize_1_1, plain, p & q,
     inference(canonicalize, [], [a])).
fof(normalize_1_2, plain, p,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_3, plain, q,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_4, plain, $false,
    inference(simplify, [],
      [normalize_1_0, normalize_1_2, normalize_1_3])).
cnf(refute_1_0, plain, ($false),
    inference(canonicalize, [], [normalize_1_4])).
\end{verbatim}
\end{frame}

\begin{frame}[fragile, label=verified-example]{Problem $p ∧ q ⊢ q ∧ p$}{Proof generated with \texttt{Athena}}
\vfill
\begin{minted}{agda}
p, q, a, goal, subgoal₀, subgoal₁ : Prop

@@

\begin{frame}[fragile, label=verified-example-2]{Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
{\footnotesize
\begin{minted}{agda}
a : Prop
a = p ∧ q

@@

proof₀ : Γ ⊢ subgoal₀
proof₀ =
  (RAA
    (atp-simplify ⊥
      (assume {Γ = Γ}
        (¬ subgoal₀))
      (atp-conjunct q
        (atp-canonicalize (p ∧ q)
          (weaken (¬ subgoal₀)
            (assume {Γ = ∅} a))))))
\end{minted}
}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-3]
  {Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
{\footnotesize
\begin{minted}{agda}
subgoal₁ : Prop
subgoal₁ = q ⇒ p

proof₁ : Γ ⊢ subgoal₁
proof₁ =
  (RAA
    (atp-simplify ⊥
      (atp-conjunct q
        (atp-canonicalize (p ∧ q)
          (weaken (¬ subgoal₁)
            (assume {Γ = ∅} a))))
      (atp-simplify ⊥
        (atp-canonicalize ((¬ p) ∧ q)
          (assume {Γ = Γ}
            (¬ subgoal₁)))
        (atp-conjunct p
          (atp-canonicalize (p ∧ q)
            (weaken (¬ subgoal₁)
              (assume {Γ = ∅} a)))))))
\end{minted}
}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-4]{Problem $p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
{\footnotesize
\begin{minted}{agda}
-- Premise.
Γ = [ a ]

@@

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal            -- q ∧ (q ⇒ p) ⇒ p
    (∧-intro proof₀ proof₁)
\end{minted}
}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug\footnote{\url{https://github.com/gilith/metis/issues/2}.} in the Printing of the Proof}
  {\texttt{Metis} v2.3 (release 20161108)}
% {}
{\footnotesize 
\begin{verbatim}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{verbatim}
}
